Reconstructing the last known movements of one of Nature's giants

### 2017-06-23 manuscript/blue-whale-ms.tex
Add CT edits to paragraph 3 main text

Baleen is formed of plates of keratin and is used by mysticete whales to filter feed. 
It is ideal for stable isotope studies because keratin grows continuously through an individual's life, and once laid down it is metabolically inert\cite{best1996stable}. 
We sampled keratin from the blue whale's baleen plate at 1cm intervals and used continuous-flow isotope ratio mass spectrometry to extract $\delta^{13}$C and $\delta^{15}$N isotope concentrations. 
Given the date of stranding (25th March 1891), and estimated baleen growth rates of 13.5cm $y^{-1}$ (see Supplemental Methods), we reconstructed a timeline for $\delta^{13}$C and $\delta^{15}$N fluctuations in the baleen over the last six years of whale's life (autumn 1885 - spring 1891). The $\delta^{13}$C and $\delta^{15}$N profiles show two distinct phases in the whale's behaviour. 
In behavioural phase one (month 1885 to May 1889), we find strong cyclical fluctuations in both $\delta^{13}$C and $\delta^{15}$N values of an approximately yearly periodicity (Fig. 1). $\delta^{13}$C and $\delta^{15}$N exhibit strong negative covariance. 
In behavioural phase two (June 1889 to March 1891) we see a sharp uptick in $\delta^{13}$C, followed by high and relatively constant $\delta^{13}$C values. ,
This is associated with an overall reduction in $\delta^{15}$N values and a weakening of the periodic $\delta^{15}$N fluctuations seen in behavioural phase one. There is also no covariance between $\delta^{13}$C and $\delta^{15}$N (Fig. 1).

>>>>

Baleen is formed of plates of keratin and is used by mysticete whales to filter feed. 
It is ideal for stable isotope studies because keratin grows continuously through an individual's life, and once laid down it is metabolically inert\cite{best1996stable}. 
We sampled keratin from the blue whale's baleen plate at 1cm intervals perpendicular to the direction of growth and used continuous-flow isotope ratio mass spectrometry determine the isotopic compositions of carbon and nitrogen (expressed as $\delta^{13}$C and $\delta^{15}$N values). 
Given the date of stranding (25th March 1891), and estimated baleen growth rates of 13.5cm $y^{-1}$ (see Supplemental Methods), we reconstructed a timeline for $\delta^{13}$C and $\delta^{15}$N fluctuations in the baleen over the last six years of whale's life (autumn 1885 - spring 1891). The $\delta^{13}$C and $\delta^{15}$N profiles show two distinct phases in the whale's behaviour. 
In behavioural phase one (month 1885 to May 1889), we find strong cyclical fluctuations in both $\delta^{13}$C and $\delta^{15}$N values with strong negative covariance (Fig. \ref{fig1}; Fig. S?). 
In behavioural phase two (June 1889 to March 1891) we see a sharp increase in $\delta^{13}$C values for approximately eight months.
This is associated with an overall reduction in $\delta^{15}$N values and a weakening of the periodic $\delta^{15}$N fluctuations seen in behavioural phase one. 
There is also no covariance between $\delta^{13}$C and $\delta^{15}$N values in behavioral phase two (Fig. \ref{fig1}).
